Racial and Ethnic Differences in ETI: Outcomes

### 2025-07-25 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
PEx models FINALLY working

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(multcomp)
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(broom)
knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
```

@@

@@

@@

@@

>>>>

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(multcomp)
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(broom)
library(emmeans)
knitr::opts_knit$set(root.dir = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
```

@@

For some reason, exacerbation models with the encounter-level data are just not working. I think it may be because there are often several pulmonary exacerbation visits very close together, so including the "Days" variable is causing problems. In the end I used a very simple logistic regression model on annualized PEx data. See `/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R` for details on making the exacerbation dataset.

@@

```{r}
# Zero-inflated models
# mod_p <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.poisson(), zi_fixed = ~1
# )
# mod_nb <- mixed_model(num_pex ~ ETI * Race,
#   data = pex_data,
#   random = ~ 1 | eDWID, zi_random = ~ 1 | eDWID,
#   family = zi.negative.binomial(), zi_fixed = ~1
# )
pex_logistic <- glmer(any_pex ~ ETI * Race + (1 | eDWID),
  data = pex_data,
  control = glmerControl(optimizer = "bobyqa"),
  family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
  gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
```

The table above shows the predicted probability of a pulmonary exacerbation pre- and post-ETI. So for example, NHW pwCF had a PEx probability of `r round(means$prob[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Non-Hispanic White"&means$ETI=="Pre-ETI"]*100,1)`)  pre-ETI and `r round(means$prob[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Non-Hispanic White"&means$ETI=="Post-ETI"]*100,1)` post-ETI. Conversely, Black pwCF had a PEx probability of `r round(means$prob[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Black or African American"&means$ETI=="Pre-ETI"]*100,1)`)  pre-ETI and `r round(means$prob[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` (95%CI: `r round(means$asymp.LCL[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` - `r round(means$asymp.UCL[means$Race=="Black or African American"&means$ETI=="Post-ETI"]*100,1)` post-ETI.

@@

```{r}
pex_logistic <- glmer(any_pex ~ ETI * Race + (1 | eDWID),
  data = pex_data[pex_data$age_eti_group == "[-Inf,18)", ],
  control = glmerControl(optimizer = "bobyqa"),
  family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
  gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
```

@@

```{r}
pex_logistic <- glmer(any_pex ~ ETI * Race + (1 | eDWID),
  data = pex_data[pex_data$age_eti_group == "[18, Inf)", ],
  control = glmerControl(optimizer = "bobyqa"),
  family = "binomial"
)
means <- data.frame(emmeans(pex_logistic, ~ Race | ETI, type = "response"))
means %>%
  gt(caption = "Probabilities by Race", rownames_to_stub = T) %>%
  fmt_number(decimals = 3)
```
